### 2017-11-27 paper/sections/1_introduction.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different domain and service.  Our exploration of this space results in a
dynamically sized cache policy that, for our initial conditions, sacrifices
negligible performance while using only 28\% of the memory required by our
hand-tuned cache.

@@

Storage systems use software-based caches to improve performance but the
policies that guide what data to evict and when to evict vary with the use
case. For example, caching file system metadata on clients and servers reduces
the number of remote procedure calls and improves the performance of
create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
patil:fast2011-giga+}. But the policies for when to evict and what data to
evict are specific to the application's behavior and the system's configuration
(hardware, settings, etc.) so a new use case may prove to be a poor match for
the selected caching
policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
variety of caching policies using a data management language/policy engine and
arrive at a customized policy that works well for our target application.  This
process of trying general policies and quickly iterating to a customized
solution shows that our prototype can adapt to different workloads.

@@

Our target application, ParSplice~\cite{perez:jctc20150parsplice}, is
representative of applications that extensively use software-based caches.
ParSplice is a molecular dynamics simulation that uses a hierarchy of caches
and a single persistent key-value store to store both observed minima across a
molecule's equation of motion (EOM) and the hundreds or thousands of partial
trajectories calculated each second during a parallel job.  The fine-grained
data annotation capabilities provided by key-value storage is a natural match
for scientific simulations like ParSplice. But these simulations rely on a
mesh-based decomposition of a physical region and result in millions or
billions of mesh cells, where each cell contains materials, pressures,
temperatures and other characteristics that are required to accurately simulate
phenomena of interest.  Unfortunately, simulations of this size saturate the
capacity and bandwidth capabilities of a single node so we need more effective
data management techniques, such as cache management or load balancing across a
cluster.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we designed a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, we find this threshold
works well for HPC applications.  For reference, a 1GB cache for a distributed
file system is too large in LANL deployments.}. ParSplice deploys a cache per
300 worker processes, so large simulations will not scale because they need
multiple caches.  Users can configure ParSplice to evict data when the cache
reaches a threshold but this solution requires tuning and parameter sweeps; the
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the expense of performance,
which is shown by the annotation to the right.  Even worse, this threshold
changes with different initial configurations and cluster setups so tuning
needs to be done for all system permutations.  Our dynamically sized cache,
shown by the thick line in Figure~\ref{fig:cache-management}, detects key
access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache with a
negligible performance degradation (within the baseline's standard deviation)
and works for a variety of initial conditions.

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
was designed for the narrow purpose of file system metadata load balancing,
this paper provides evidence that the approach is more broad.  Specifically,
the collection of abstractions in Mantle provides a general control plane that
improves the performance of metadata access. So in this paper we refer to
Mantle as a policy engine that injects policies written in our data management
language directly into a running service, such as a file system or key-value
store.  Rather than co-designing a policy directly into the service, which
requires domain-specific knowledge to find and change policies, we expose the
policies in a general way so even developers unfamiliar with the domain can
quickly deploy solutions.  We show that our framework:

@@

This last contribution is explored in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}, where we try a range of policies from different
disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
that the collection of policies we designed for cache management in ParSplice
are very similar to the policies used to load balance metadata in the Ceph file
system (CephFS) suggesting that there is potential for automatically adapting
and generating policies dynamically.

>>>>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different domain and service.  Our exploration of this space results in a
dynamically sized cache policy that sacrifices negligible performance while
using 32-66\% less memory than the default ParSplice configuration.

@@

Storage systems use software-based caches to improve performance but the
policies that guide what data to evict and when to evict vary with the use
case. For example, caching file system metadata on clients and servers reduces
the number of remote procedure calls and improves the performance of
create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
patil:fast2011-giga+}. But the policies for what data to evict and when to
evict are specific to the application's behavior and the system's configuration
(hardware, settings, etc.) so a new use case may prove to be a poor match for
the selected caching
policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
variety of caching policies using our data management language/policy engine and
arrive at a customized policy that works well for our target application.  This
process of trying general policies and quickly iterating to a customized
solution shows that our prototype can adapt to different workloads.

@@

Our target application, ParSplice~\cite{perez:jctc20150parsplice}, is
representative of applications that extensively use software-based caches.
ParSplice is a molecular dynamics simulation that uses a hierarchy of caches
and a single persistent key-value store to store both observed minima across a
molecule's equation of motion (EOM) and the hundreds or thousands of partial
trajectories calculated each second during a parallel job.  The fine-grained
data annotation capabilities provided by key-value storage is a natural match
for scientific simulations like ParSplice. But these simulations rely on a
mesh-based decomposition of a physical region and result in millions or
billions of mesh cells, where each cell contains materials, pressures,
temperatures and other characteristics that are required to accurately simulate
phenomena of interest.  Unfortunately, simulations of this size saturate the
capacity and bandwidth capabilities of a single node so we need more effective
data management techniques.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we design a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, we find this threshold
works well for HPC applications.  For reference, a 1GB cache for a distributed
file system is too large in LANL deployments.}. Furthermore, ParSplice deploys
a cache per 300 worker processes, so large simulations need more caches and
will use even more memory.  Users can configure ParSplice to evict data when
the cache reaches a threshold but this solution requires tuning and parameter
sweeps; the ``Cache (too small)" curve in Figure~\ref{fig:cache-management}
shows how a poorly configured cache can save memory but at the cost of
performance, which is shown by the text annotation to the right.  Even worse, this
threshold changes with different initial configurations and cluster setups so
tuning needs to be done for all system permutations.  Our dynamically sized
cache, shown by the thick line in Figure~\ref{fig:cache-management}, detects
key access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache with a
negligible performance degradation (within the baseline's standard deviation)
and works for a variety of initial conditions.

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
was designed for the narrow purpose of file system metadata load balancing,
this paper provides evidence that the approach is more broad.  Specifically,
the collection of abstractions in Mantle provide a general control plane that
improves the performance of metadata access. So in this paper we refer to
Mantle as a policy engine that injects policies written in our data management
language directly into a running service, such as a file system or key-value
store.  Rather than co-designing a policy directly into the service, which
requires domain-specific knowledge to find and change policies, we expose the
policies in a general way so even developers unfamiliar with the domain can
quickly deploy solutions.  We show that our framework:

@@

This last contribution is explored in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}, where we try a range of policies from different
disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
that the collection of policies we design for cache management in ParSplice
are very similar to the policies used to load balance metadata in the Ceph file
system (CephFS) suggesting that there is potential for automatically adapting
and generating policies dynamically.

### 2017-11-27 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

The computation can be parallelized by adding more workers or by adding tasks
to parallelize individual workers.  The workers are stateless and read initial
coordinates from the data store each time they begin generating segments. Since
worker tasks do not maintain their own history, they can end up reading the
same coordinates repeatedly. To mitigate the consequences of these repeated
reads, ParSplice provisions a hierarchy of processes to act as caches that sit
in front of a single persistent database.  Values are written to each tier and
reads traverse up the hierarchy until they find the data.  Caches also reside
on the workers to service reads/writes from its tasks.

@@

Our evaluation uses the total ``trajectory length" as the goodness metric. This
value is the duration of the overall trajectory produced by ParSplice. At
ideal efficiency, the trajectory length should increase with the square root of
the wall-clock time, since the wall-clock cost of time-stepping the system by
one simulation time unit increases in proportion of the total number of atoms.

@@

\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (text annotations) and request load (bars) after a one hour run
with a different number of tasks(\(x\) axis). While the keyspace size and
capacity is relatively modest the memory usage scales linearly with the number
of tasks. This is a problem if we want to scale to Trinitite's 3000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 20GB for the cache.  This memory utilization easily eclipses
the 3\% memory usage per node threshold we set earlier, even without factoring
in the usage from other workers.

@@

Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  The \(\Delta_1\) growth rate
adds atoms every 100K microseconds while the \(\Delta_2\) growth rate adds
atoms every 1 million microseconds. So \(\Delta_2\) has a smaller growth rate
resulting in hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on
the \(y2\) axis).  Values smaller than \(\Delta_2\)'s growth rate or a
temperature of 400 degrees results in very little database activity because
state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
growth rate or a temperature of 4000 degrees result in an equally meaningless
simulation as transitions are unrealistic.  This figure demonstrates that small
changes to \(\Delta\) can have a strong effect on the timing and frequency with
which new EOM minima are discovered and referenced. This finding suggests that
we need a flexible policy language and engine to explore these trade-offs.

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run.  The growth rate, temperature, and number of
workers have a predictable effect on the structure of the keyspace.
Figure~\ref{fig:methodology-keys} shows that the number of reads changes with
different initial conditions (\(\Delta\)), but that the spatial locality of key
accesses is similar ({\it e.g.}, some keys are still \(5\times\) more popular
than others).  Figure~\ref{fig:motivation-regimes} shows how entropy for
different growth rates has temporal locality, as the reads per second for
\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
along the time axis.  Trends also exist for temperature and number of workers
but are omitted here for space. This structure means that we can learn the key
access patterns and adapt the storage system accordingly (this motivates
Section~\S\ref{sec:dom-specific}).

>>>>

The computation can be parallelized by adding more workers or by adding tasks
to parallelize individual workers.  The workers are stateless and read initial
coordinates from the data store each time they begin generating segments. Since
worker tasks do not maintain their own history, they can end up reading the
same coordinates repeatedly. To mitigate the consequences of these repeated
reads, ParSplice provisions a hierarchy of caches that sit in front of a single
persistent database.  Values are written to each tier and reads traverse up the
hierarchy until they find the data.

@@

@@

\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (text annotations) and request load (bars) after a one hour run
with a different number of tasks(\(x\) axis). While the keyspace size and
capacity is relatively modest the memory usage scales linearly with the number
of tasks. This is a problem if we want to scale to Trinitite's 3000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 8GB for one cache.  This memory utilization easily eclipses
the 3\% memory usage per node threshold we set earlier, even without factoring
in the usage from other workers.

@@

Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  The \(\Delta_1\) growth rate
adds atoms every 100K microseconds while the \(\Delta_2\) growth rate adds
atoms every 1 million microseconds. So \(\Delta_2\) has a smaller growth rate
resulting in hotter keys and a smaller keyspace.  Values smaller than
\(\Delta_2\)'s growth rate or a temperature of 400 degrees results in very
little database activity because state transitions take too long. Similarly,
values larger than \(\Delta_1\)'s growth rate or a temperature of 4000 degrees
result in an equally meaningless simulation as transitions are unrealistic.
This figure demonstrates that small changes to \(\Delta\) can have a strong
effect on the timing and frequency with which new EOM minima are discovered and
referenced. This finding suggests that we need a flexible policy language and
engine to explore these trade-offs.

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) look like
the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the key access patterns and
adapt the storage system accordingly (this motivates
Section~\S\ref{sec:dom-specific}).

### 2017-11-27 paper/sections/3_methodology.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%  \item background: CephFS implementation
To explore software-defined cache management, we use the data management
language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
prototype in that paper was called Mantle and it was originally built on CephFS
so administrators could control file system metadata data management policies.
We now refer to Mantle as a policy engine that supports our data management
language.  The basic premise is that data management policies can be expressed
with a simple API consisting of ``when", ``where", and ``how much". ``when"
controls how aggressive or conservative the decisions are; ``where" controls
how distributed or concentrated the data should be; and ``how much" controls
the amount of data that should be sent. There is also a ``load" policy that
lets administrators specify how to collapse many metrics into a single load
metric ({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).

@@

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a service.  Administrators write policies from whatever domain they
choose ({e.g.}, statistics, machine learning, storage system) and the policies
are embedded into the runtime by Mantle. When the service makes decisions it
executes the administrator-defined policies for when/where/how much and returns
a decision.  To do this, the service needs to be modified to (1) provide an
environment of metrics and (2) identify where policies are set. These
modification points are shown by the colored boxes in Figure~\ref{fig:mantle}
and described below.

Administrators write policies in
Lua that are executed whenever the service needs to make a data management
decision. We chose Lua for simplicity, performance, and portability; it is a
scripting language with simple syntax, which allows administrators to focus on
the policies themselves; it was designed as an embeddable language, so it is
lightweight and does less type checking; and it interfaces nicely with C/C++.

@@

\subsubsection{Environment of Metrics} services expose cluster metrics for
describing resource utilization and time series metrics for describing accesses
to some data structure over time. Table~\ref{table:metrics} shows how these
metrics are accessed from the policies written by administrators.

For cluster metrics, the service passes a dictionary to Mantle, indexed by
server name and metric. Policies access the cluster metric values by indexing
into a Lua table using \texttt{server} and \texttt{metric}, where server is a
node identifier ({e.g.}, MPI Rank, metadata server name) and metric is a
resource name.  For example, a policy written for an MPI-based service can
access the CPU utilization of the first rank in a communication group using:

@@

Other examples metrics used for file system metadata load balancing are shown
by the ``environment" box in Figure~\ref{fig:mantle}. The measurements and
exchange of metrics between servers is done by the service; Mantle in CephFS
leverages metrics from other servers collected using CephFS's heartbeats.

For time series metrics, the service passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses. For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
ts = array()
for i=1,arraysize() do
  for time,value in string.gmatch(ts, (%w+=%w)) do
    if value == 'mydirectory' then
      count = count + 1
    end
  end
end
\end{minted}
\end{figure}

The service uses a pointer to the time series to facilitate time series with
many values, like accesses to a database or directory in the file system
namespace. This decision limits the time series metrics to only include values
from the {\it current } node, although this is not a limitation of Mantle
itself.

@@

% env of metrics
Using Mantle cluster metrics, we expose cache size, CPU utilization, and memory
pressure of the worker node to the cache management policies. In
Section~\S\ref{sec:parsplice-keyspace-analysis} we only end up using the cache
size although the other metrics proved to be valuable debugging tools. Using
Mantle time series metrics, we expose accesses to the cache as a list of
\texttt{timestamp, key} pairs. In Section~\S\ref{sec:dom-specific}, we
explore a key access pattern detection algorithm that uses this metric.

>>>>

%  \item background: CephFS implementation
To explore software-defined cache management, we use the data management
language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
prototype in that paper was called Mantle and it was originally built on CephFS
so administrators could control file system metadata load balancing policies.
We now refer to Mantle as a policy engine that supports our data management
language.  The basic premise is that data management policies can be expressed
with a simple API consisting of ``when", ``where", and ``how much". The ``when" policy
controls how aggressive or conservative the decisions are; ``where" controls
how distributed or concentrated the data should be; and ``how much" controls
the amount of data that should be sent. There is also a ``load" policy that
lets administrators specify how to collapse many metrics into a single load
metric ({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).

@@

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a service.  Administrators write policies in Lua from whatever
domain they choose ({\it e.g.}, statistics, machine learning, storage system)
and the policies are embedded into the runtime by Mantle.  We chose Lua for
simplicity, performance, and portability; it is a scripting language with
simple syntax, which allows administrators to focus on the policies themselves;
it was designed as an embeddable language, so it is lightweight and does less
type checking; and it interfaces nicely with C/C++.  When the service makes
decisions it executes the administrator-defined policies for when/where/how
much and returns a decision.  To do this, the service needs to be modified to
(1) provide an environment of metrics and (2) identify where policies are set.
These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

@@

\subsubsection{Environment of Metrics} services expose \textbf{cluster} metrics
for describing resource utilization and \textbf{time series} metrics for
describing accesses to some data structure over time. Table~\ref{table:metrics}
shows how these metrics are accessed from the policies written by
administrators.

For cluster metrics, the service passes a dictionary to Mantle. Policies access
the cluster metric values by indexing into a Lua table using \texttt{server}
and \texttt{metric}, where server is a node identifier ({\it e.g.}, MPI Rank,
metadata server name) and metric is a resource name.  Metrics used for file
system metadata load balancing are shown by the ``environment" box in
Figure~\ref{fig:mantle}. The measurements and exchange of metrics between
servers is done by the service; Mantle in CephFS leverages metrics from other
servers collected using CephFS's heartbeats.  For example, a policy written for
an MPI-based service can access the CPU utilization of the first rank in a
communication group using:

@@

For time series metrics, the service passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses.  The
service uses a pointer to the time series to facilitate time series with many
values, like accesses to a database or directory in the file system namespace.
This decision limits the time series metrics to only include values from the
{\it current } node, although this is not a limitation of Mantle itself.  For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
d = timeseries()       -- get data (d) from service
for i=1,d:size() do    -- iterate over timeseries
  ts, value = d:get(i) -- index into timeseries 
  if value == 'mydirectory' then
    count = count + 1
  end
end
\end{minted}
\end{figure}

@@

% env of metrics
Using Mantle cluster metrics, we expose cache size, CPU utilization, and memory
pressure of the worker node to the cache management policies. In
Section~\S\ref{sec:arch-specific} we only end up using the cache size although
the other metrics proved to be valuable debugging tools. Using Mantle time
series metrics, we expose accesses to the cache as a list of \texttt{timestamp,
key} pairs. In Section~\S\ref{sec:dom-specific}, we explore a key access
pattern detection algorithm that uses this metric.
